Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex

Stack safety properties need to describe the behavior of machine code. but they naturally
talk about function activations and stack contents---abstractions that
are (typically) not visible at machine level. To bridge this gap,
our properties are defined in terms of a {\em security semantics} layered on top of
the standard execution semantics of the machine.  This semantics identifies certain
state transitions of the machine as {\em security-relevant operations}, which update
a notional {\em security context}.  This context consists of
an (abstract) stack of function activations, each associated with a {\em view}
that maps each element of the machine state (memory location or register)
to a {\em security class} (active, sealed, etc.) which specifies how the activation
can access the element.
The action of the security-relevant operations on the context is defined by a set
of rules that depend on how the machine code makes use of the stack and registers
to implement the function abstraction.

We define stack safety properties in
terms of traces of {\em observable events} rather than internal machine states.
This is particularly important for
specifying lazy enforcement, which works because the deferred check catches problems
before they can affect the trace.
Events might include certain function calls (i.e., system calls performing I/O)
or writes to special addresses representing memory-mapped regions, etc.
%The nature of events is a parameter of the framework, and can be specialized
%to any notion of observable behavior.

Finally, we define our high-level security properties---integrity, confidentiality,
and well-bracketed control flow---in terms of the security classes of state
elements. These take the form of predicates on states and contexts that use variants,
assertions about future return states, and comparison of event traces to capture the
desired behavior without reference to internal implementation details.
We introduce these properties by example in \cref{sec:example}. In the remainder
of this section we describe the underlying framework in more detail.

Our security semantics has similarities to the overlay semantics proposed
by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it does not restrict
the behavior of the underlying machine in any way. Rather, it tracks additional context
about the history of security-relevant operations, which inform the criterion a machine
must satisfy in order to correctly obey the properties.\apt{Sharpen this comparison?}

>>>>

Stack safety properties need to describe the behavior of machine code. but they naturally
talk about function activations and stack contents---abstractions that
are typically not visible at machine level. To bridge this gap,
our properties are defined in terms of a {\em security semantics} layered on top of
the standard execution semantics of the machine.  The security semantics identifies certain
state transitions of the machine as {\em security-relevant
  operations}, which update
a notional {\em security context}.  This context consists of
an (abstract) stack of function activations, each associated with a {\em view}
that maps each element of the machine state (memory location or register)
to a {\em security class} (active, sealed, etc.) specifying how the activation
can access the element.
The action of a security-relevant operation on the context is defined by a set
of rules that describe how the machine code makes use of the stack and registers
to implement the function abstraction.

We define stack safety properties in
terms of traces of {\em observable events} rather than internal machine states.
This is particularly important for
specifying lazy enforcement, which catches bad behaviors after they occur but
before they can affect the trace.
Observable events might include specific function calls (e.g., system
calls that perform I/O), writes to special addresses representing
memory-mapped regions, etc.
%The nature of events is a parameter of the framework, and can be specialized
%to any notion of observable behavior.

Finally, we define high-level security properties---integrity, confidentiality,
and well-bracketed control flow---in terms of the security classes of state
elements. These take the form of predicates on states and contexts
that use variants\bcp{undefined},
assertions about future return states, and comparison of event traces to capture the
desired behavior without reference to internal implementation details.
We introduce these properties by example in \cref{sec:example} and
formally in \cref{sec:formal}. In the remainder
of this section we describe the underlying framework in more detail.

Our security semantics has similarities to the overlay semantics
proposed by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it
does not restrict the behavior of the underlying machine in any
way\bcp{Will people not familiar with Skorstengaard understand
  that?}. Rather, it tracks additional context about the history of
security-relevant operations, which informs the criterion a machine
must satisfy in order to correctly obey the properties.\apt{Sharpen
  this comparison?}\bcp{+1.  Or maybe, in the interest of time and to
  keeping this section flowing, move it to related work.}
